MLISA: Machine Learning-based \\ Intelligent Service Agents

### 2014-11-14 papers/webservice-coalition-learning/webservice_coalition_learning.tex
changes

Now we want to generate a set of feature vectors representing the communities of web services, for our training purposes. For communities we add two additional external features as discussed in section \ref{s:ef} that reflect the community's estimation of its quality metrics compared to other web services. Therefore we formulate a \emph{Community Feature Vector (CFV)} as $CFV_{<C>} = \{f_1,...f_5\}$ having community of $k$ web services. $C = {ws_1,...ws_k}$. The features $f_1$ through $f_5$ represent the \emph{execution time}, \emph{throughput}, \emph{availability}, the \emph{external parameters 1 and 2} correspondingly. A set of communities, with their feature vectors and utilities evaluated provides our algorithm with raw training data set. We call this set of communities the \emph{template vector} and the set of feature vectors associated with the \emph{template vector} would be the \emph{community feature vector set (CFVS)}.

\subsubsection{feature engineering}\label{sss:feng}
Let $CFVS = \{C_1, C_2,..., C_N\}$ be the community feature vector set with $N$ communities. Based on $CFVS$ set we create a $|N \times N|$ gain matrix for $t$ different time slots. Each entry of $gain_{n,m}^{t}$corresponds to utility gain of web services in $C_n$, when $C_n$ joins $C_m$, based on community and web service data available for time $t$. All these values have to be evaluated by implementing the community with members of $ws_i \in C_n$ and $ws_j \in C_m$ and evaluating its utility (see Equation \ref{u_c_load}). Therefore $gain_{n,m}^{t} = U_{C_n \cup C_m}^{t} - U_{C_{n}}^{t}$. Evaluating the utility gain for all entries of $gain$ matrix is can be a costly process with very big number of $N$, therefore size of feature vector $|N|$ should be chosen carefully.

Now we let our set of communities in $CFVS$ set, within $|T|$ time frame iterations, choose the the best communities to join. Each community is provided with the corresponding row of data from the $gain$ matrix. Basicly $C_i$ is provided with the data in row $i$ of matrix $gain_{N,N}$ which is all the possible utility values it can gain by joining different communities. By ordering the list, each community is utilized with an ordered set of preferences over other communities it likes to join. We define $>_{i}^t$ as preference order of community $i$ at time $t$.

Let $P_{C_i}^t = \{C_1 >_{i}^t C_2 >_{i}^t .... >_{i}^t C_n\}$ be an ordered set of preferences for community $C_i$ at time $t$. Based on this ordered set, we define $K^t(C_i, k)$ as set of $k$ most preferred communities of community $i$ at time $t$.
%\subsubsection{feature vector generation}\label{sss:fvg}
\begin{equation}\label{h_t_pref_top}
\begin{split}				
K^t(C_i, k) = &\Big\{C_x | C_x >_{i}^t C_y; \forall C_y \in CFVS\ \\
				      &\wedge C_y \notin \{K^t(i, 1),K^t(i, 2),...,K^t(i, k-1)\} \Big\}				
\end{split}
\end{equation}
Based on $K^t(C_i, k)$ we define a set of communities for $C_i$ in which they are k most preferred communities for community $i$ and also community $i$ belongs to k most preferred communities of any of communities in set $K^t(C_i, k)$. This basically implies the preference is both sided.
\begin{equation}\label{l_t_top_both}
\begin{split}	
L^t(C_i,k) = &\Big\{C_j | C_j \in K^t(C_i, k) \\
             &\wedge C_i \in \bigcup_{C_x \in K^t(C_j, k)}K^t(C_x, k)\Big\}
\end{split}
\end{equation}

Having these sets built based on training model, we are able to build a decision tree where each community's preferences are set as well as the joined events that took place in different time slots. In addition to best choices, web services have access to ordered other choices and can look for the second best or third best if their first try is rejected by the target community. The consequences of each try level is analyzed in more details in the following section where we launch experiments and investigate how effective is the use of decision tree with different decision layers in joining other communities and enhancing the overall utility.

@@

Our goal is to create a decision making profile for each community in $CFVS$ set. We creating an environment where they can experience the outcome of difference strategies. The result will be a decision tree of the possible feasible and utility increasing moves over time.

We let communities pick best communities maximizing their utilities over different time frames. At time $t = 1$, we let each community in the $CFVS$ set choose the best community which is single community in set $C_j = K^t(C_i, 1)$. If community $C_j$ also has the highest preference to join $C_i$ meaning $L^t(C_i, k=1)$ set is not empty, they would join together. Having set $k = 1$ is a very strict and hardly satisfiable. In order to relax the requirement the value of $k$ is being increased as $t$ increases by a rate $r$. Basically $k = r \times t$, On early steps of simulation web services are more strict and as time goes on we let them choose, second, third best options too. However increasing $k$ fast increases the time complexity of $K^t(C_i, k)$ by $O(k^2 \times n.log(n) + n.log(n))$.

When communities $C_i$ and $C_j$ are in each others top $k$ preference set, their combination is being added to the list of possible communities that can join others at time $t+1$. Also for each community $C_i$ in our initial $CFVS$ set, we maintain a tree having the community $C_i$ as root. Its children are all the communities $C_i$ decided to join. As scenario progresses over time, the merged communities may decide to join other communities. When community $C_i$ and $C_j$ decides to join each other and create community $C_k$, the new community $C_k$ will be added as child to both $C_i$ and $C_j$ nodes. At the end of process, each community is utilized with a tree of set of possible combination of communities it can join.

Having created $|n|$ trees, for each community $C_i$, our communities are utilized with the different possible paths they can take to maximize their utilities. Using a distance function, communities and web services can utilize these results and find the $C_i$ which closely resembles their parameters. They will be utilized with a tree which has the strategies the community can take to join our communities which will be best possible path to take and communities to join while most likely being accepted by other communities since its a two sided preference.

\section{Experiments}\label{s:experiments}
We have implemented MLISA in Java\footnote{Source code of implementation and data is available at: \emph{https://github.com/Marooned202/MLISA}}. We have extracted the set of features for 4532 web services in 64 different time slots, through a data set provided by \cite{10.1109/ISSRE.2011.17}. By choosing 86 web services out of these set, and choosing a subset of set of all possible combinations of size 2,3,4 of these 86 web services, we have chosen 10,000 communities and evaluated the feature vectors and utilities they can have in 64 different time slots. This provides us with the initial training feature set of size $|CFVS| = 10,000$ communities. Based on equation \ref{u_c_load} the utility of these communities are estimate and the gain matrix $|10000 \times~ 10000|$ of all possible ways of merging these 10,000 communities, is generated\footnote {The template vector and gain matrix generated are available at \emph{https://github.com/Marooned202/MLISA/tree/master/wsds/data/run}} and each community has the ordered preference among other communities known in the set.

We let communities and web services adopt their strategies based on our $MLISA$ decision making mechanism. Each community based on the decisions adopted will generate with a decision tree profile. We let MLISA run four times with different $r$ rates of 0.05, 0.07, 0.10, 0.20. With slow rate of $r = 0.05$ we increase $k$ in equation \ref{u_c_normal} every 20 time frames which will happen only three times in our 64 step experiment and in the case of $r = 0.20$ k is increased much faster, once every 5 time frames, which increased the complexity of $L^t(C_i,k)$ search for each community in $CFVS$ set.

Figure \ref{utility_gain_value} depicts utility gain value the 10 random communities in each of the four runs. The utility gain is the increase of utility the communities gain by cooperating and joining other communities. Comparing the different search rates, we can see that the high $r$ values of $0.20$ and $0.10$, all 10 samples have been able to find and collaborate with communities which gives them the best values. However for $r=0.7$ in 1 cases, and $r=0.5$ in 2 cases they have not been able to find the optimal community to form. Figure \ref{utility_gain_ratio} is the same experiment illustrating the ratio of utility gain compared to the communities' initial utility value, rather than the abstract gain value.

>>>>

Now, we want to generate a set of feature vectors representing the communities of web services for training purposes. For communities, we add two additional external features, as discussed in Section \ref{s:ef} that reflect the community's estimation of its quality metrics compared to other web services. Therefore, we formulate a \emph{Community Feature Vector (CFV)} as $CFV_{<C>} = \{f_1,...f_5\}$ having a community of $k$ web services. $C = {ws_1,...ws_k}$. The features $f_1$ through $f_5$ represent the \emph{execution time}, \emph{throughput}, \emph{availability} and the \emph{external parameters 1 and 2} correspondingly. A set of communities, with their feature vectors and utilities evaluated, provides our algorithm with a raw training data set. We call this set of communities the \emph{template vector}, and the set of feature vectors associated with the \emph{template vector} is referred to as the \emph{community feature vector set (CFVS)}.

\subsubsection{feature engineering}\label{sss:feng}
Let $CFVS = \{C_1, C_2,..., C_N\}$ be the community feature vector set with $N$ communities. Based on the $CFVS$ set, we create an $|N \times N|$ gain matrix for $t$ different time slots. Each entry of $gain_{n,m}^{t}$corresponds to a utility gain of web services in $C_n$, when $C_n$ joins $C_m$, based on community and web service data available for time $t$. All of these values have to be evaluated by implementing the community with members of $ws_i \in C_n$ and $ws_j \in C_m$ and evaluating its utility (see Equation \ref{u_c_load}). Therefore $gain_{n,m}^{t} = U_{C_n \cup C_m}^{t} - U_{C_{n}}^{t}$. Evaluating the utility gain for all entries of the $gain$ matrix can be a costly process with a very big number of $N$, size of the feature vector $|N|$ should be chosen carefully.

Now, we let our set of communities in the $CFVS$ set, within $|T|$ time frame iterations, choose the the best communities to join. Each community is provided with the corresponding row of data from the $gain$ matrix. Basically, $C_i$ is provided with the data in row $i$ of matrix $gain_{N,N}$, which is all of the possible utility values it can gain by joining different communities. By ordering the list, each community is utilized with an ordered set of preferences over other communities it likes to join. We define $>_{i}^t$ as the preference order of community $i$ at time $t$.

Let $P_{C_i}^t = \{C_1 >_{i}^t C_2 >_{i}^t .... >_{i}^t C_n\}$ be an ordered set of preferences for community $C_i$ at time $t$. Based on this ordered set, we define $K^t(C_i, k)$ as a set of the $k$ most preferred communities of community $i$ at time $t$.
%\subsubsection{feature vector generation}\label{sss:fvg}
\begin{equation}\label{h_t_pref_top}
\begin{split}				
K^t(C_i, k) = &\Big\{C_x | C_x >_{i}^t C_y; \forall C_y \in CFVS\ \\
				      &\wedge C_y \notin \{K^t(i, 1),K^t(i, 2),...,K^t(i, k-1)\} \Big\}				
\end{split}
\end{equation}
Based on $K^t(C_i, k)$, we define a set of communities for $C_i$ in which they are the k most preferred communities for community $i$ and also community $i$ belongs to the k most preferred communities out of all of the communities in set $K^t(C_i, k)$. This basically implies the preference is both-sided.
\begin{equation}\label{l_t_top_both}
\begin{split}	
L^t(C_i,k) = &\Big\{C_j | C_j \in K^t(C_i, k) \\
             &\wedge C_i \in \bigcup_{C_x \in K^t(C_j, k)}K^t(C_x, k)\Big\}
\end{split}
\end{equation}

Having these sets built based on the training model, we are able to build a decision tree wherein each community's preferences are set, as well as the joined events that took place in different time slots. In addition to best choices, web services have access to other ordered choices and can look for the second best or third best if their first try is rejected by the target community. The consequences of each try level is analyzed in more detail in the following section, in which we launch experiments and investigate the effectiveness of the use of a decision tree with different decision layers in joining other communities and enhancing the overall utility.

@@

Our goal is to create a decision making profile for each community in the $CFVS$ set. We are creating an environment where they can experience the outcomes of different strategies. The result will be a decision tree of the feasible and utility increasing moves over time.

We let communities pick the best communities for maximizing their utilities over different time frames. At time $t = 1$, we let each community in the $CFVS$ set choose the best community, which is a single community in set $C_j = K^t(C_i, 1)$. If community $C_j$ also has the highest preference to join $C_i$ meaning the $L^t(C_i, k=1)$ set is not empty, they would join together. Having set $k = 1$ is a very strict and hardly satisfiable condition. In order to relax the requirement, the value of $k$ is being increased by a rate $r$ proportional to time slot $t$. Basically, $k = r \times t$. On early steps of simulation web services and communities are more strict but as time goes on, we let them choose second and then third best options too. However, increasing $k$ quickly increases the time complexity of $K^t(C_i, k)$ by $O(k^2 \times n.log(n) + n.log(n))$.

When communities $C_i$ and $C_j$ are in each other's top $k$ preference set, their combination is added to the list of possible communities that can join others at time $t+1$. Also, for each community $C_i$ in our initial $CFVS$ set, we maintain a tree with the community $C_i$ as its root. Its children are all of the communities that $C_i$ decided to join. As the scenario progresses over time, the merged communities may decide to join other communities. When community $C_i$ and $C_j$ decide to join each other and create community $C_k$, the new community $C_k$ will be added as a child to both $C_i$ and $C_j$ nodes. At the end of the process, each community is utilized with a tree of set of possible combination of communities it can join.

Having created $|n|$ trees, for each community $C_i$, our communities are utilized with the different possible paths they can take to maximize their utilities. Using a distance function, communities and web services can utilize these results and find the $C_i$ that closely resembles their parameters. They will be utilized with a tree that has the strategies the community can use to join our communities which will be best possible path to take and communities to join while most likely being accepted by other communities since its a two sided preference.

\section{Experiments}\label{s:experiments}
We implemented MLISA in Java\footnote{Source code of implementation and data is available at: \emph{https://github.com/Marooned202/MLISA}}. We have extracted the set of features for 4532 web services in 64 different time slots through a data set provided by \cite{10.1109/ISSRE.2011.17}. By choosing 86 web services out of these sets and choosing a subset of all possible combinations of sizes 2, 3, and 4 of these 86 web services, we have chosen 10,000 communities and evaluated the feature vectors and utilities they can have in 64 different time slots. This provides us with the initial training feature set of size $|CFVS| = 10,000$ communities. Based on Equation \ref{u_c_load} the utility of these communities are estimated, and the gain matrix $|10000 \times~ 10000|$ of all possible ways of merging these 10,000 communities, is generated\footnote {The template vector and gain matrix generated are available at \emph{https://github.com/Marooned202/MLISA/tree/master/wsds/data/run}}, and each community has the ordered preference among other communities known in the set.

We let communities and web services adopt their strategies based on our $MLISA$ decision making mechanism. Based on the decisions adopted, each community will generate a decision tree profile. We let MLISA run four times with different $r$ rates of 0.05, 0.07, 0.10 and 0.20. With the slow rate of $r = 0.05$, we increase $k$ in Equation \ref{u_c_normal} for every 20 time frames, which will happen only three times in our 64-step experiment. In the case of $r = 0.20$, k increases much faster, at a rate of once every 5 time frames, which increases the complexity of the $L^t(C_i,k)$ search for each community in the $CFVS$ set.

Figure \ref{utility_gain_value} depicts the utility gain value the 10 random communities in each of the four runs. The utility gain is the increase of utility the communities gain by cooperating and joining other communities. Comparing the different search rates, we can see that the high $r$ values of $0.20$ and $0.10$, all 10 samples have been able to find and collaborate with communities which gives them the best values. However for $r=0.7$ in 1 cases, and $r=0.5$ in 2 cases they have not been able to find the optimal community to form. Figure \ref{utility_gain_ratio} is the same experiment illustrating the ratio of utility gain compared to the communities' initial utility value, rather than the abstract gain value.
